---
title: 'Business Statistics: helpful R functions'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Binomial distribution

**Given**: number of independent trials ($k$), and a probability of success ($p$) on each trial

We typically want to know the probability that the number of successes (let's refer to this as $X$) is equal to some value, or range of values. 

**Example**: In 10 flips of a biased coin where the probability of *Heads* is 0.6, what is the probability of obtaining less than or equal to 3 Heads.

**Solution**: We want: $P(X=0) + P(X=1) + P(X=2)+ P(X=3)$. In the code chunk below, this is computed using a manual method and using an r formula: <code>dbinom</code>.

```{r}

# the arguments to dbinom are (x, number of trials, probability of success), where x 
# you want to find the probability that X = x
sum ( dbinom(0:3, 10, p = 0.7) ) # 0.01

# and manually using the binomial probability function
choose(10,0) * 0.7 ^ 0 * (1-0.7)^10 + choose(10,1) * 0.7 ^ 1 * (1-0.7)^9 + choose(10,2) * 0.7 ^ 2 * (1-0.7)^8 + choose(10,3) * 0.7 ^ 3 * (1-0.7)^7

```


### Normal distribution

**Given**: the mean ($\mu$) and standard deviation ($\sigma$) of a random variable ($X$) that follows a normal distribution

We typically want to know the probability that $X$ is in some range. This is achieved with the <code>pnorm</code> function.

**Example**: The distribution of cat weights is normal with mean = 9 and standard deviation = 0.9. What is the probability that$\ldots$

```{r}

pnorm(10, mean = 9, sd = 0.9) # P(X < 10)

pnorm(10, mean = 9, sd = 0.9) - pnorm(8, mean = 9, sd = 0.9) # P(8 < X < 10)

```

We may also be interested in finding a certain percentile of a distribution. This is achieved with the <code>qnorm</code> function. 

For example, what is the 80% percentile of the cat weight distribution?

```{r}
qnorm(0.8, mean = 9, sd = 0.9) # 80th percentile of cat weights
```

### Poisson Distribution

**Given**: The average rate of occurrence ($\lambda$) for some event within a given time window

We typically want to know the probability that some number of events occur within a unit of time. The probability function for the poisson distribution is used by the function <code>dpois</code>.

**Example**: On a certain island, there are 9 earthquakes on average that occur during the course of a year. What is the probability that more than 10 earthquakes are observed in a year?

**Solution**: Let's refer to the count of earthquakes as $X$. Then, the solution is to find: $1 - P(X \leq 10)$

```{r}
1 - sum( dpois(0:10, lambda = 9) ) 
```









